---
title: "Examples: drawit()"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Examples: drawit()}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message=FALSE}
library(youdrawitR)
```

```{r, include = FALSE, message=FALSE}
library(ggplot2)
library(dplyr)
library(palmerpenguins)
library(tidyverse)
```


## Introduction

`drawit()` is designed to mimic the New York Times “You Draw It” tool, where viewers are shown part of a graph and asked to predict the trend before the full data is revealed. Earlier versions of the `youdrawitR` package required users to work with JavaScript and D3. This version only requires familiarity with ggplot2.

In this vignette, we will demonstrate the `drawit()` tool under a variety of circumstances, using both real and simulated data. Although this package was originally created for graphical testing, it can be used more broadly in any setting where users want to sketch or draw on a graph. 

### How to Use `drawit()`

- First, create a `ggplot2` object using any combination of up to two of the following geoms:
  - geom_point
  - geom_smooth
  - geom_line
- Then, pass that plot into `drawit()`, either with a pipeline or by saving it as an object first
- The output will display a graph with a yellow box. Click inside the box and move from left to right to draw your prediction. As you move, a line will appear. The yellow box indicates the section of the graph that still needs to be completed, so make sure to draw across the full highlighted region.

## Examples

### Real Data Examples

Because `youdrawitR` builds on `ggplot2`, it works naturally with real datasets. This makes it useful when you want users to estimate or continue a trend they see between variables.

Note: `drawit() `works best with smaller datasets because it tries to match each x-value in the original data. This is especially helpful if you want to collect the user-drawn values and merge them back with the plotted data. If you need more flexibility, see `sketchit()`.

Let's see it used with `mtcars`.

```{r}
(ggplot(data = mtcars, aes(x = wt, y = mpg)) +
  geom_point()) |> 
  drawit()
```

The smoothness, or jitteriness, of the drawn path depends on how densely packed the data are. Compare the `mtcars` example to the `palmerpenguins` dataset.  

```{r}
penguins_adelie <- dplyr::filter(penguins, species == "Adelie")

(ggplot(data = penguins_adelie, aes(x = bill_length_mm, y = bill_depth_mm)) +
  geom_point()) |> 
  drawit()
```

In this example, the drawn path is much more jittery and requires more careful mouse movement to connect smoothly. Moving your mouse slowly across the plot often helps the function capture the movement more accurately. If jitteriness becomes a problem, consider reducing the density of the data or increasing the height and width when rendering. This often makes the interaction easier to control.

If you need more flexibility, check out `sketchit()`. That function is more freeform and can be especially useful when the data are dense.

While `drawit()` is not a perfect replica of `ggplot2`, it does preserve much of the styling information from the `ggplot` object you pass into it.

```{r}
(ggplot(data = mtcars, aes(x = wt, y = mpg)) +
    geom_line(linewidth = 2, color = "red") +
    labs(title = "Drawit Cars Example",
         subtitle = "Hope You Enjoy!",
         x = "Weight",
         y = "Miles Per Gallon") +
   scale_y_continuous(limits = c(0, 32))+
   scale_x_continuous(limits = c(1.5, 7))
 ) |> 
  drawit()
```

### Simulated Data Examples

`drawit()` is especially useful with simulated data, since it was originally designed for graphical testing. The function includes several options that allow you to adjust the graph to fit your use case.

```{r, message = FALSE}
# Here is the simulated data we will be using
data = tibble(x = seq(1, 25, .5), 
              y = exp((x-15)/30),
              ypoints = exp(((x-15)/30) + rnorm(30, 0, 0.1)))

```

Often, it is helpful to ask users to continue the trend beyond the observed data. You can control where the drawing region begins with the `draw_start` argument in `drawit()`.

```{r}
(ggplot(data, aes(x, ypoints)) +
  geom_point(data = data |>  mutate(ypoints2 = ifelse(x < 10, ypoints, NA)),
             aes(y = ypoints2),
             size = 2, 
             colour = "magenta"
             ) +
  scale_y_continuous(limits = c(0.5, 1.5)) +
  labs(x = "X-Label", y = "Y-Label")) |>
  drawit(draw_start = 10)
```

`drawit()` supports up to two geoms at a time. For example, you can include the true line up to the starting point so the user has a guide before continuing the pattern.

```{r}
(ggplot(data, aes(x, ypoints)) +
  geom_point(data = data |>  mutate(ypoints2 = ifelse(x < 10, ypoints, NA)),
             aes(y = ypoints2),
             size = 2, 
             colour = "magenta"
             ) +
   geom_line(data = data |>  mutate(y = ifelse(x < 10, y, NA)),
            aes(y = y)) +
  scale_y_continuous(limits = c(0.5, 1.5)) +
  labs(x = "X-Label", y = "Y-Label")) |>
  drawit(draw_start = 10)
```

Alternatively, you can reveal the full line after the user finishes drawing by using the `show_on_finish` argument. If `show_on_finish` is set to `TRUE`, `drawit()` will use the secondary plot as the "revealed" plot.

```{r}
(ggplot(data, aes(x, ypoints)) +
  geom_point(data = data |>  mutate(ypoints2 = ifelse(x < 10, ypoints, NA)),
             aes(y = ypoints2),
             size = 2, 
             colour = "magenta"
             ) +
   geom_line(data = data,
            aes(y = y)) +
  scale_y_continuous(limits = c(0.5, 1.5)) +
  labs(x = "X-Label", y = "Y-Label")) |>
  drawit(draw_start = 10, show_on_finish = TRUE)
```

### Shiny Integration

The `drawit()` function integrates directly with `Shiny`, enabling interactive drawing inputs to be captured and reused within reactive workflows. When used inside a `Shiny` application, `drawit()` behaves differently than in static contexts. Instead of returning only an `r2d3` widget, it returns a list, which contains:

- `youdrawit_plot`: the interactive drawing widget
- `points`: a reactive tibble containing the user’s drawn data

The `points` object is a `reactive()` expression that resolves to a `tibble` with:

- `x`: original (as well as some interpolated) x-values
- `y`: corresponding user-drawn values

Because `drawit()` enforces a continuous, one-to-one drawing, this output represents a fully specified function over the draw space. 

::: {.callout-note}
To enable communication between the browser and the Shiny server, you must supply the `shiny_message_loc` argument when using `drawit()` in Shiny. This value must match the input ID used internally for message passing. If this argument is missing or empty, `drawit()` will throw an error.
:::
